Coding the LBG Wind (in 1D)

### 2010-08-28 Winds/papers/Theory/First/Notes/lbg_clumps.tex
blah
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@5153 ac11df5e-a788-c23b-a8e1-c758c323924d

\begin{center}
{\Large Simple Exploration of the LBG Clump Model (Steidel+10)}
\end{center}

@@

\begin{my_itemize}
\item \nc: Number of clumps in a given shell of width $\Delta r$
\item $d_c$: Size of the clump (assumed to be cubes)
\item $\sigma$: Cross-section of the clump ($d_c^2$)
\item $n_c$: Number density of a clump (Hydrogen)
\item $m_c$: Mass of a clump, i.e. $m_p n_c d_c^3$
\item $A$: Total area of clumps, i.e. $\mnc \sigma$
\item $f_c$: Covering fraction of clumps at radius $r$
\end{my_itemize}

@@

\begin{equation}
f_c(r) = f_{c,max} \ltp \frac{r}{r_{\rm inner}} \rtp^{-\gamma} \cmma
\label{eqn:covering}
\end{equation}

with $\gamma \approx 0.5$ and velocity

@@

Consider a shell of width $\Delta r$ at radius $r$.  The surface area
of this shell is $4\pi r^2$.  To achieve a covering fraction $f_c$,
one requires

\begin{equation}
\mnc(r) = \frac{4 \pi r^2}{\sigma} \cdot f_c(r) = 
\frac{4 \pi f_{c,max}}{\sigma \, (\rm kpc^{1/2})} \cdot r^{3/2}
\end{equation}
clumps within the shell.  It is evident that the number of clumps
increases rather steeply with radius.  [Note: we could allow for
varying clump size by doing the calculation in terms of $A$ but I
don't think it would matter much.]

Now consider the mass profile of this clumpy wind.  Again, assuming
identical clumps at all radii, we have the mass per shell of

\begin{equation}
M_{\Delta r} = m_c \mnc 
\end{equation}
Normalizing to the mass in the first shell ($r = 1 - 2$\,kpc), we get
the cumulative mass profile shown in Figure~\ref{fig:mass}.
We find that there is 10000$\times$ more mass in the wind beyond that
in the innermost shell!

\clearpage

\begin{figure}[ht]
\begin{center}
\includegraphics[width=4.5in,angle=180]{../Figures/fig_lbg_clump_mass.pdf}
\end{center}
  \caption{Cumulative mass profile, normalized to the mass within
    $r=1-2$\,kpc, for identical clumps.  There is over 10000$\times$
    more mass beyond 2\,kpc than within that first shell. 
}
  \label{fig:mass}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=4.5in,angle=180]{../Figures/fig_lbg_clump_eandp.pdf}
\end{center}
  \caption{Cumulative energy and momentum profiles.
}
  \label{fig:eandp}
\end{figure}

Figure~\ref{fig:eandp} shows the energy and momentum
profiles, again relative to that in the inner most shell assuming
$E_{\Delta r} = M_{\Delta r} v^2$ and $P = M_{\Delta r} v$.  The
results are similar to the mass profile because the velocity curve
rises so steeply at small radii.

\vskip 0.2in

Now consider an estimation of the absolute value for the mass.  I'll
begin by assuming a size for the clump of $d_c = 100$\,pc.  If we
desire this clump to be optically thick for SiII~1526, then we need a
Si$^+$ number density $n_{\rm Si^+} = 10^{14} \cm{-2} / d_c$.  Lastly,
I will assume solar abundance, i.e. $n_H = 10^{4.5} n_{\rm Si^+}$.
This gives a number density $n_H \approx 0.01 \cm{-3}$.  I think it
might be hard to have much lower number density. Each clump,
then, has a mass $m_c = 253 \msol$. To
get a 80\%\ covering fraction at 2\,kpc, we will need $\mnc \approx
4000$ clumps or a total of $\approx 10^6 \msol$.  Finally, we estimate
that the total mass in the wind is $10^{10} \msol$ with $\sim 90\%$ of
that mass beyond 30\,kpc.  All of this material is travelling at
800\,\kms.

\vskip 0.1in

So, is this at all plausible?  And, if not, can one monkey with the
size of my clumps etc.\ to make a plausible scenario?

>>>>

\begin{center}
{\Large Simple Exploration of the LBG Clump Model (Steidel+10): v2}
\end{center}

@@

\begin{my_itemize}
\item \nc: Number of clumps in a given shell of width $\Delta r$
\item $\Delta v_c$: Internal velocity dispersion of the clump (e.g.\
  10 km/s)
\item $\Delta r$: Width of the shell. This will relate to the internal
  velocity dispersion of the clumps.
\item $n$: Number density of clumps at radius $r$
\item $d_c$: Size of the clump (assumed to be cubes)
\item $\sigma$: Cross-section of the clump ($d_c^2$)
\item $n_c$: Number density of a clump (Hydrogen)
\item $m_c$: Mass of a clump, i.e. $m_p n_c d_c^3$
%\item $A$: Total area of clumps in $\Delta r$, i.e. $\mnc \sigma$
\item $f_c$: Covering fraction of clumps at radius $r$
\end{my_itemize}

@@

\begin{equation}
f_c(r) = f_{c,max} \ltp \frac{r}{r_{\rm inner}} \rtp^{-\gamma} \cmma
\label{eqn:covering}
\end{equation}
with $\gamma \approx 0.5$ and velocity

@@

Consider a shell of width $\Delta r$ at radius $r$.   The number of
clumps in this shell is:

\begin{equation}
\mnc = n \cdot \Delta r \cdot 4 \pi r^2
\end{equation}
And the covering fraction will be

\begin{equation}
f_c = \frac{\sigma \mnc}{4 \pi r^2} = n \sigma \Delta r
\end{equation}
Lastly, the thickness of a given shell relates to the overlap of the
clumps in velocity space.  Given an internal velocity dispersion $\Delta
v_c$ for each clump, we have:

\begin{equation}
\Delta r = \Delta v_c \ltp \frac{dv}{dr} \rtp^{-1}  \perd
\end{equation}
This starts to look like the Sobolev equations, but not quite.  The
key issue here is that $\Delta r$ varies with radius.  Because the LBG
velocity curve flattens, $\Delta r$ increases quite rapidly with $r$.

If we like, we can rewrite these equations in terms of the density of
the clumps

\begin{equation}
n(r) = \frac{f_c(r)}{\sigma} \frac{1}{\Delta r} \cmma
\end{equation}
the mass of a shell of $\Delta r$

\begin{equation}
M_{\Delta r} = \frac{f}{\sigma} \cdot 4 \pi r^2 \cdot m_c
\end{equation}
We see (as in v1 of this write-up) that the mass in each shell goes as
$r^{3/2}$ given the functional form of the covering fraction.  But the
mistake I had made is that the thickness of the shells ($\Delta r$) is
increasing with radius.  So the cumulative mass doesn't increase as
rapidly.

Now consider an estimation of the absolute value for the mass.  I'll
begin by assuming a size for the clump of $d_c = 100$\,pc.  If we
desire this clump to be optically thick for SiII~1526, then we need a
Si$^+$ number density $n_{\rm Si^+} = 10^{14} \cm{-2} / d_c$.  Lastly,
I will assume solar abundance, i.e. $n_H = 10^{4.5} n_{\rm Si^+}$.
This gives a number density $n_H \approx 0.01 \cm{-3}$.  I think it
might be hard to have much lower number density. Each clump,
then, has a mass $m_c = 253 \msol$.

With this clump size and mass, we get the density, mass, and flux
profiles shown in Figure~\ref{fig:aclump}.  Each of these curves start
at $r = 2$\,kpc because at smaller radii the clumps overlap.  Again,
the mass approaches $10^{10} \msol$ and most of the mass lies at the
outer radii.  The clump density drops very rapidly with radius such
that the clump flux is nearly constant.  I think this is by design
given the shells have thickness $\Delta r \propto (dv/dr)^{-1}$, but
I'm not sure.

So, is this at all plausible?  And, if not, can one monkey with the
size or $\Delta v_c$ of my clumps etc.\ to make a plausible scenario?

\begin{figure}[ht]
\begin{center}
\includegraphics[width=5.5in]{../Figures/fig_lbg_aclump.pdf}
\end{center}
  \caption{$\Delta r$, density, mass, and flux profiles for the LBG
    wind model assuming $\Delta v_c = 10 \mkms$ and that the clump has
    $d_c = 100$\,pc and $m_c = 250 \msol$.
}
  \label{fig:aclump}
\end{figure}
